---
title: "pumpR Workout Report"
author: "Magnus Nissel"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book

--- 
output: bookdown::gitbook
documentclass: book
bibliography: [pumpr.bib]
biblio-style: apalike
link-citations: yes
description: "Creating a report from workout logs"
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

# Overview
This is a year-by-year overview generated from my workout logs ^[I used to track via [Fitocracy](https://www.fitocracy.com) and now use [FitNotes](https://play.google.com/store/apps/details?id=com.github.jamesgay.fitnotes)]. I have excluded cardio activities (e.g. indoor rowing, swimming) and there are a couple of untracked workouts missing from the data. 

## Workout frequency
The first plot shows the total number of days on which I logged a workout, the other displays the monthly sum of sets to account for workout length. 

```{r warning=FALSE, cache=FALSE, echo=FALSE, fig.width=8}
source("plot.R")

# Full workout log
log_df = read.csv(PLOG_PATH)
log_df$MONTH <- month(log_df$DATE, label=T)
log_df$YEAR <- year(log_df$DATE)
# Workout days
days_df <- log_df %>% distinct(DATE)
days_df$MONTH <- month(days_df$DATE, label=T)
days_df$YEAR <- year(days_df$DATE)
# Plot heatmaps
days_ym <- group_by(days_df, YEAR, MONTH)
workout_days <- summarise(days_ym, days=n())
day_fig <- heat_plotter(workout_days, "MONTH", "YEAR", "days", "Workout days", "Workouts per month by year")

log_ym <- group_by(log_df, YEAR, MONTH)
sets <- summarise(log_ym, Sets= n(), num_reps=mean(REPS))
set_fig <- heat_plotter(sets, "MONTH", "YEAR", "Sets", "Number of sets", "Sets per month by year")

day_fig
set_fig
```

My most active year was cleary 2013, with 20+ workouts per month from March to July. However, these were generally shorter workouts with body part splits. I am not quite
sure about the three-month break in 2014, but my least active period was from September 2015 to April 2016. I tried some CrossFit in December - which I did no bother logging - and switched gyms ^[My previous membership ended in January and the new gym opened in May], but most of the absence was due work keeping me more busy than usual.

```{r warning=FALSE, cache=FALSE, echo=FALSE, fig.width=8}
workouts_months <- days_df %>% group_by(MONTH) %>% summarise(Workouts =  n())
p <- ggplot(workouts_months, aes(x = MONTH, y=Workouts, fill=Workouts)) + geom_bar(stat = "identity") 
p <- p + theme_tufte(ticks=FALSE) + ggtitle("Workouts per month (all years)") + xlab("") + ylab("") 
p <- p + scale_fill_gradient2(low='white', high='steelblue')  + theme(legend.position="none")
p

sets_months <- log_df %>% group_by(MONTH) %>% summarise(Sets =  n())
p <- ggplot(sets_months, aes(x = MONTH, y=Sets, fill=Sets)) + geom_bar(stat = "identity") 
p <- p + ggtitle("Sets per month (all years)") + xlab("") + ylab("") + theme_tufte(ticks=FALSE)
p <- p + scale_fill_gradient2(low='white', high='steelblue')  + theme(legend.position="none")
p

```

Apparently I am *really* motivated in August.

## The _Big Four_

  - **Squat**
  - **Deadlift**
  - **Bench Press**
  - **Overhead Press**

### Traditional ^[_Flat Barbell Bench Press_, _Barbell (Sumo) Deadlift_, _(High/Low Bar) Barbell Back Squat_, _Standing Barbell Overhead Press_]

```{r warning=FALSE, cache=FALSE, echo=FALSE, out.width="100%"}

four <- filter(log_df, BIG_FOUR != "Other")
four <- filter(four, !grepl("\\*", BIG_FOUR))
four$YEAR <- year(four$DATE)
four <- four  %>% group_by(YEAR, BIG_FOUR) %>% dplyr::summarise(Sets = n(), Reps = sum(REPS), Volume = sum(VOLUME))

p <- ggplot(four, aes(x = BIG_FOUR, y=Volume, fill=Volume)) + geom_bar(stat = "identity") 
p <- p + ggtitle("Combined volume") + xlab("") + ylab("") + theme_tufte(ticks=FALSE)
p <- p + scale_fill_gradient2(low='white', high='steelblue') + facet_wrap( ~ YEAR, ncol=2) + theme(legend.position="none")
p

d_four <- filter(log_df, BIG_FOUR != "Other")
d_four <- filter(d_four, !grepl("\\*", BIG_FOUR))
d_four <- d_four  %>% group_by(DATE, BIG_FOUR) %>% dplyr::summarise(Sets = n(), Reps = sum(REPS), Volume = sum(VOLUME))

p <- ggplot(d_four, aes(x=as.Date(DATE), y=Sets, group=BIG_FOUR, color=BIG_FOUR)) + geom_line()
p <- p + scale_x_date() + xlab("") + ylab("") + theme_tufte(ticks=FALSE) 
p <- p + ggtitle("Sets over time") + facet_grid(BIG_FOUR ~ .) + theme(legend.position="none")
p

```

### With variations ^[All kinds of variants from _Machine Hack Squats_ to _One-Arm Dumbbell Presses_ are considered here.]


```{r warning=FALSE, cache=FALSE, echo=FALSE, out.width="100%"}
four <- filter(log_df, grepl("\\*", BIG_FOUR))
four <- four  %>% group_by(BIG_FOUR) %>% dplyr::summarise(Sets = n(), Reps = sum(REPS), Volume = sum(VOLUME))



p <- ggplot(four, aes(x = BIG_FOUR, y=Volume, fill=Volume)) + geom_bar(stat = "identity") 
p <- p + ggtitle("Combined volume") + xlab("") + ylab("") + theme_tufte(ticks=FALSE)
p <- p + scale_fill_gradient2(low='white', high='steelblue')  + theme(legend.position="none")
p
```

## Implements

__Note__: Not all exercises have been assigned the correct implement, yet. 

```{r warning=FALSE, cache=FALSE, echo=FALSE, out.width="100%"}

four <- log_df  %>% group_by(IMPLEMENT) %>% dplyr::summarise(Sets = n(), Reps = sum(REPS), Volume = sum(VOLUME))



p <- ggplot(four, aes(x = IMPLEMENT, y=Sets)) + geom_bar(stat = "identity", colour="steelblue", fill="steelblue") 
p <- p + ggtitle("Number of sets") + xlab("") + ylab("") + coord_flip() + theme_tufte(ticks=FALSE)
p <- p + theme(legend.position="none") 
p



```

## Movements

__Note__: Not all exercises have been assigned the correct movement type, yet. 

```{r warning=FALSE, cache=FALSE, echo=FALSE, out.width="100%"}

four <- log_df  %>% group_by(MOVEMENT) %>% dplyr::summarise(Sets = n(), Reps = sum(REPS), Volume = sum(VOLUME))



p <- ggplot(four, aes(x = MOVEMENT, y=Sets, fill=Sets)) + geom_bar(stat = "identity") 
p <- p + ggtitle("Number of sets") + xlab("") + ylab("") + coord_flip() + theme_tufte(ticks=FALSE)
p <- p + theme(legend.position="none") + scale_fill_gradient2(low='white', high='steelblue') 
p
```
