---
title: "pumpR Workout Report"
author: "Magnus Nissel"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book

--- 
output: bookdown::gitbook
documentclass: book
bibliography: [project.bib]
biblio-style: apalike
link-citations: yes
description: "A linguistic investigation of the discourse on basketball on the Web"
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

# Overview
This is a year-by-year overview generated from my workout logs ^[I used to track via Fitocracy and now use FitNotes]. I have excluded cardio activities (e.g. indoor rowing, swimming) and there are a couple of untracked workouts missing from the data. 

## Frequency
The first plot shows the total number of days on which I logged a workout, the other displays the monthly sum of sets to account for workout length. 

```{r warning=FALSE, cache=FALSE, echo=FALSE, fig.width=8}
source("plot.R")

# Full workout log
log_df = read.csv(LOG_PATH)
log_df$MONTH <- month(log_df$DATE, label=T)
log_df$YEAR <- year(log_df$DATE)
# Workout days
days_df <- log_df %>% distinct(DATE)
days_df$MONTH <- month(days_df$DATE, label=T)
days_df$YEAR <- year(days_df$DATE)
# Plot heatmaps
days_ym <- group_by(days_df, YEAR, MONTH)
workout_days <- summarise(days_ym, days=n())
day_fig <- heat_plotter(workout_days, "MONTH", "YEAR", "days", "Workout days", "Workouts per month by year")

log_ym <- group_by(log_df, YEAR, MONTH)
sets <- summarise(log_ym, num_sets= n(), num_reps=mean(REPS))
set_fig <- heat_plotter(sets, "MONTH", "YEAR", "num_sets", "Number of sets", "Sets per month by year")

day_fig
set_fig
```

My most active year was cleary 2013, with 20+ workouts per month from March to July. However, these were generally shorter workouts with body part splits. I am not quite
sure about the three-month break in 2014, but my least active period was from September 2015 to April 2016. I tried some CrossFit in December - which I did no bother logging - and switched gyms ^[My previous contract ended in January and the new gym opened in May], but most of the absence was due work keeping me more busy than usual and the flu.

```{r warning=FALSE, cache=FALSE, echo=FALSE, fig.width=8}
workouts_months <- days_df %>% group_by(MONTH) %>% summarise(workouts =  n())
ggplot(workouts_months, aes(x = MONTH, y=workouts)) + geom_bar(stat = "identity") + ggtitle("Workouts per month (all years)") + xlab("") + ylab("") + theme_tufte()

sets_months <- log_df %>% group_by(MONTH) %>% summarise(num_sets =  n())
ggplot(sets_months, aes(x = MONTH, y=num_sets)) + geom_bar(stat = "identity")  + ggtitle("Sets per month (all years)") + xlab("") + ylab("") + theme_tufte()


```

Apparently I am *really* motivated in August.



